Exploring mathematical objects from custom-tailored mathematical universes

### 2022-04-06 paper-filmat.tex
typo (thanks to Lukas!)
typo

The number~$M(X)$ computed by~$M$ depends on the input/output behaviour of~$X$,
the chosen ordering of infinite-time Turing machines, and on details of the
interleaving simulation and the comparison procedure -- but it does not depend on
the implementation of~$X$ or on its specific choice of rational
approximations~$X(n)$. The search terminates since there is at least one
infinite-time Turing machine which represents the same real number as~$X$ does,
namely~$X$ itself.}

@@

Sheaf toposes take up a special place in the history of topos theory: If the
base~$X$ is allowed to be a \emph{site} instead of a topological space, the
resulting toposes constitute the large class of Grothendieck toposes, the
original notion of toposes. Categorically, the passage from topological spaces
to sites is rather small, but the resulting increase in flexibility is
substantial and fundamental to modern algebraic geometry.

@@

\subsection{The Zariski topos}
\label{sect:the-zariski-topos}
The starting point is the observation that
while the field~$\RR$ of ordinary real numbers does not contain infinitesimals
(except for zero), the ring~$\RR[\varepsilon]/(\varepsilon^2)$ of \emph{dual
numbers} does. This ring has the cartesian product~$\RR \times \RR$ as its
underlying set and the ring operations are defined such
that~$\varepsilon^2 = 0$, where~$\varepsilon \defeq \langle 0,1 \rangle$:
\[
  \langle a,b \rangle + \langle a',b' \rangle \defeq
  \langle a+a', b+b' \rangle \qquad
  \langle a,b \rangle \cdot \langle a',b' \rangle \defeq
  \langle aa', ab'+a'b \rangle
\]
We write~$\langle a,b \rangle$ more clearly
as~$a+b\varepsilon$.

>>>>

The number~$M(X)$ computed by~$M$ depends on the input/output behaviour of~$X$,
the chosen ordering of infinite-time Turing machines, and on details of the
interleaving simulation and the comparison procedure -- but it does not depend on
the implementation of~$X$ nor on its specific choice of rational
approximations~$X(n)$. The search terminates since there is at least one
infinite-time Turing machine which represents the same real number as~$X$ does,
namely~$X$ itself.}

@@

Sheaf toposes take up a special place in the history of topos theory: If the
base~$X$ is allowed to be a \emph{site} instead of a topological space, the
resulting toposes constitute the large class of Grothendieck toposes, the
original notion of toposes. Categorically, the step from topological spaces
to sites is rather small, but the resulting increase in flexibility is
substantial and fundamental to modern algebraic geometry.

@@

\subsection{The Zariski topos}
\label{sect:the-zariski-topos}
The starting point is the observation that
while the field~$\RR$ of ordinary real numbers does not contain infinitesimals
(except for zero), the ring~$\RR[\varepsilon]/(\varepsilon^2)$ of \emph{dual
numbers} does. This ring has the cartesian product~$\RR \times \RR$ as its
underlying set and the ring operations are defined such
that~$\varepsilon^2 = 0$, where~$\varepsilon \defeq \langle 0,1 \rangle$:
\[
  \langle a,b \rangle + \langle a',b' \rangle \defeq
  \langle a+a', b+b' \rangle \qquad
  \langle a,b \rangle \cdot \langle a',b' \rangle \defeq
  \langle aa', ab'+a'b \rangle
\]
We write~$\langle a,b \rangle$ more suggestively
as~$a+b\varepsilon$.
